Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-10-27 short.tex
bunches of changes

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as an unordered \emph{trace},\footnote{For our domain we do not require the trace to be ordered -- so it is similar to, but not exactly the same as, the execution trace of the program.} since it corresponds to the commands that a program would have issued but lacks the high-level structure that determines how the program decided to issue them.
%For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.
The second step involves synthesizing a high-level program capable of producing the trace identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \cite{solar2008program}. The program synthesizer can search the space of possible programs for one capable of producing the desired trace -- inducing structures like
symmetries, loops, or conditional branches.
Although these program do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

\noindent \textbf{The Trace Hypothesis.} The set of commands issued by a program,
which we call a \emph{trace},
is the correct liaison between high-dimensional unstructured perceptual input and high-level structured symbolic representations.

The roadmap of our paper is structured around the trace hypothesis, as outlined in Fig.~\ref{roadmap}.
\begin{figure}[h]
  \begin{tikzpicture}
    \node[ thick,anchor = west,inner sep=0pt,label=below:{\small Perceptual input}](observation) at (0,0) {\includegraphics[width = 1.5cm]{figures/expert-39-trimmed.png}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceSource) at (3,0.5){    \begin{lstlisting}[basicstyle = \scriptsize\ttfamily]
line, line,
rectangle,
line, ...
\end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceImage) at (3.22,-0.5) {
      \includegraphics[width = 0.9cm]{figures/39-parse.png}}; 
    \node(trace)[draw,thin,fit = (traceImage) (traceSource), label = below:{{\small Trace}}] {};

@@

    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([yshift = -15,xshift = -115]trace.south) -- node[below = 6pt] {Section~\ref{neuralNetworkSection}: Perception$\to$Trace} ([yshift = -15,xshift = -5]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([xshift = 5,yshift = -15]trace.south) -- node[below = 6pt] {Section~\ref{programSynthesisSection}: Trace$\to$Program} ([xshift = 140,yshift = -15]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
([xshift = 150,yshift = -15]trace.south) -- node[below = 6pt] {Section~\ref{applicationsSection}: Applications} ([xshift = 250,yshift = -15]trace.south);
  \end{tikzpicture}
  \caption{Both the paper and the system pipeline are structured around the \emph{trace hypothesis}}\label{roadmap}
\end{figure}

Prior work addresses converting hand drawings to vectors~\cite{huang2017shape,Nishida:2016:ISU:2897824.2925951} and converting synthetic images to symbolic parses~\cite{nsd,DBLP:journals/corr/Beltramelli17}.
Our new contributions are: (1) A model that converts sketches to high-level programs. (2) An algorithm for learning a policy for efficiently searching for programs. (3) The trace hypothesis: a framework for going from perception to programs.
  \end{itemize}

@@

\section{Neural architecture for inferring drawing commands}\label{neuralNetworkSection}

We developed a deep network architecture for efficiently inferring a
trace, $T$, from an image, $I$.  Our model constructs the
trace one drawing command at a time.  When predicting the next drawing
command, the network takes as input the target image $I$ as well as
the rendered output of previous drawing commands.  Intuitively, the
network looks at the image it wants to explain, as well as what it has
already drawn.  It then decides either to stop drawing or proposes
another drawing command to add to the trace; if it decides
to continue drawing, the predicted primitive is rendered to its
``canvas'' and the process repeats.

Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Tbl.~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{A token is an atomic symbol  -- so the model is \emph{not} predicting lines of code character-by-character, but instead knows about the syntactic structure of primitive commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. For
example, the network first decides to emit the \verb|circle| token
conditioned on the image features, then it emits the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicts the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. We constrain
coordinates to lie on a discrete $16\times 16$ grid, and later explain
how to remove this restriction.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\cite{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

\tikzset{>=latex}
\begin{figure}[h]
  \begin{tikzpicture}
  \node[draw,blue,ultra thick,anchor = west,inner sep=0pt,label=below:Target image: $I$](observation) at (0,-1) {\includegraphics[width = 2cm]{figures/expert-18.png}};
  \node[draw,blue,thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 2cm,label=below:Canvas: render$(T)$] (canvas) at (0,-4) {};
  \draw[lightgray,ultra thin,step = 0.125] ([xshift = 0.5,yshift = 0.5]canvas.south west) grid ([xshift = -0.5,yshift = -0.5]canvas.north east);
  % draw partial image on canvas
  \draw (0.375cm,-3.25cm) circle (0.125cm);
  \draw (0.625cm,-3.625) circle (0.125cm);
  \draw (0.75cm,-3.375) circle (0.125cm);

@@

  \node(next)[draw,very thick,fit = (l1) (l2) (l3), dashed, label = below:{Next line of code}] {};

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the trace of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\cite{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[h]
  \caption{Primitive drawing commands currently supported by our model.}
\label{drawingCommandTable}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes drawing command inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\end{table}

For the model in Fig.~\ref{architecture}, the distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta \left[t_k | a_\theta \left(f_\theta(I,\text{render}(T)) | \{t_j\}_{j = 1}^{k - 1}\right), \{t_j\}_{j = 1}^{k - 1}\right]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is a trace, $\theta$ are the
parameters of the neural network, $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network), and $a_\theta(\cdot|\cdot)$ is an attention mechanism. The distribution over
traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of trace $T$, the subscripts
on $T$ index drawing commands within the trace (so $T_n$ is a sequence of tokens: $t_1t_2\cdots t_K$), and the \verb|STOP|
token is emitted by the network to signal that the trace
explains the image.

We train the network by sampling traces $T$ and target
images $I$ for randomly generated scenes
and maximizing
Eq.~\ref{objective} with respect to $\theta$ by gradient ascent.
We trained the network on $10^5$ scenes, which takes a little less than a day on an Nvidia TitanX GPU.

Training does not require backpropagation across the entire sequence
of drawing commands: drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.  In a
sense, this model is like an autoregressive variant of
Attend-Infer-Repeat~\cite{eslami1603attend}, but critically, \emph{without a learned
  recurrent memory state}, e.g. an RNN.  We can do without an RNN
because we have access to ground truth (image, trace) pairs.  This
allows us to handle scenes with many objects without
worrying about the conditioning of gradients as they propagate
over long sequences, and also makes training more straightforward (it
is just maximum likelihood estimation of the model parameters).

Our network suffices to ``derender'' synthetic images like those shown in
Fig.~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathbb{R}$ uses the distance between two images as a proxy for a
likelihood. Unconventionally, the target distribution of the SMC sampler
includes the likelihood under the proposal distribution.
Intuitively, both the proposal
distribution and the distance function offer complementary signals for
whether a drawing command is correct.
%Empirically, accuracy is much better when we combine these signals.

Fig.~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We compare with a natural baseline that models the problem as a kind of image captioning (LSTM in Fig.~\ref{syntheticResults}).
Given the target image, this baseline produces the trace in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

\begin{figure}[H]\centering
  \begin{minipage}{6cm}
    \begin{minipage}[t]{3cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-58-7.png}}\end{minipage}%
    \begin{minipage}[t]{3cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-33-5.png}}\end{minipage}
    \caption{Network is trained to infer traces for randomly generated scenes like the two shown above. See supplement for details of the training data generation.}\label{trainingData}
  \end{minipage}\hfill%
  \begin{minipage}{7.5cm}\centering
    \includegraphics[width = 7cm]{figures/editDistance36.png}
%    \includegraphics[width = 4cm]{figures/accuracy.png}
      \caption{Parsing~\LaTeX~output after training on diagrams with $\leq 12$ objects. It generalizes to scenes with many more objects. Neither the stochastic search nor the neural network are sufficient on their own. \# particles varies by model: we compare the models \emph{with equal runtime} ($\approx 1$ sec/object)}\label{syntheticResults}
    \end{minipage}

  \end{figure}

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Fig.~\ref{handDrawingExamples}; see supplement for details).

\begin{wrapfigure}{R}{6cm}
\centering  \begin{minipage}[t]{2.5cm}\centering\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}\\
   \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): a hand drawing. (b): Rendering of the trace our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) for a noisy rendering of (b).}\label{handDrawingExamples}
\end{wrapfigure}

Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
\begin{multline}
-\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx\\ |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
\end{multline}
%Intuitively,
%$L_{\text{learned}}(\cdot|\cdot)$ approximates the distance between
%the trace we want and the trace we have so far.
Pixel-wise distance
metrics are sensitive to the details of how
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details.
So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We drew 100 figures by hand; see Fig.~\ref{lotsOfHandDrawings}.
These were drawn carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid, 
we made the drawings on graph paper.

For each drawing we annotated a ground truth trace and asked the neurally guided SMC sampler
to produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because the model sometimes makes mistakes % on hand drawings,
we envision it working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

>>>>

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as a \emph{trace set} since it corresponds to the set of commands in a program's execution trace but lacks the high-level structure that determines how the program decided to issue them.
%For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.
The second step involves synthesizing a high-level program capable of producing the trace set identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \citep{solar2008program}. The program synthesizer searches the space of possible programs for one capable of producing the desired trace -- inducing structures like
symmetries, loops, or conditional branches.
Although these program synthesizers do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

\noindent \textbf{The Trace Hypothesis.} The set of commands issued by a program,
which we call a \emph{trace set},
is the correct liaison between high-dimensional unstructured perceptual input and high-level structured symbolic representations.

The roadmap of our paper is structured around the trace hypothesis, as outlined in Fig.~\ref{roadmap}.
\begin{figure}[h]
  \begin{tikzpicture}
    \node[ thick,anchor = west,inner sep=0pt,label=below:{\small Perceptual input}](observation) at (0,0) {\includegraphics[width = 1.5cm]{figures/expert-39-trimmed.png}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceSource) at (3,0.5){    \begin{lstlisting}[basicstyle = \scriptsize\ttfamily]
line, line,
rectangle,
line, ...
\end{lstlisting}};
    \node[ultra thick,anchor = west,inner sep=0pt](traceImage) at (3.3,-0.5) {
      \includegraphics[width = 0.9cm]{figures/39-parse.png}}; 
    \node(trace)[draw,thin,fit = (traceImage) (traceSource), label = below:{{\small Trace set}}] {};

@@

    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([yshift = -15,xshift = -115]trace.south) -- node[below = 6pt] {{\small Section~\ref{neuralNetworkSection}: Perception$\to$Trace set}} ([yshift = -15,xshift = -5]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
    ([xshift = 5,yshift = -15]trace.south) -- node[below = 6pt] {{\small Section~\ref{programSynthesisSection}: Trace Set$\to$Program}} ([xshift = 140,yshift = -15]trace.south);
    \draw[decoration = {brace,mirror,raise = 5pt},decorate,thick]
([xshift = 150,yshift = -15]trace.south) -- node[below = 6pt] {{\small Section~\ref{applicationsSection}: Applications}} ([xshift = 250,yshift = -15]trace.south);
  \end{tikzpicture}
  \caption{Both the paper and the system pipeline are structured around the \emph{trace hypothesis}}\label{roadmap}
\end{figure}

The new contributions of this work are:
(1) A model that converts sketches to high-level programs:
in contrast to converting images to vectors or low-level parses~\citep{huang2017shape,Nishida:2016:ISU:2897824.2925951,nsd,DBLP:journals/corr/Beltramelli17,im2latex}.
(2) An algorithm for learning a policy for efficiently searching for programs,
building on Levin search~\citep{levin1973universal}
and generalizing recent work like DeepCoder~\citep{BalGauBroetal16}.
(3) The trace hypothesis: a framework for going from perception to programs,
which connects this work to other models of program induction that use traces,
like the  Neural Program Interpreter~\citep{DBLP:journals/corr/ReedF15}.

@@

\section{Neural architecture for inferring trace sets}\label{neuralNetworkSection}

We developed a deep network architecture for efficiently inferring a
trace set, $T$, from an image, $I$.
Our model is a combination of  ideas from 
Neurally-Guided Procedural Models~\citep{ritchie2016neurally}
and Attend-Infer-Repeat~\citep{eslami1603attend}.
Our model constructs the
trace one drawing command at a time.  When predicting the next drawing
command, the network takes as input the target image $I$ as well as
the rendered output of previous drawing commands.  Intuitively, the
network looks at the image it wants to explain, as well as what it has
already drawn.  It then decides either to stop drawing or proposes
another drawing command to add to the trace; if it decides
to continue drawing, the predicted primitive is rendered to its
``canvas'' and the process repeats.

Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Tbl.~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{A token is an atomic symbol  -- so the model is \emph{not} predicting lines of code character-by-character, but instead knows about the syntactic structure of primitive commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. In Fig.~\ref{architecture} the network first decided to emit the \verb|circle| token
conditioned on the image features, then it emitted the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicted the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. We currently constrain
coordinates to lie on a discrete $16\times 16$ grid,
but the grid could be made arbitrarily fine.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\citep{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

\tikzset{>=latex}
\begin{figure}[t]
  \begin{tikzpicture}
  \node[draw,blue,ultra thick,anchor = west,inner sep=0pt,label=below:Target image: $I$](observation) at (0,-1) {\includegraphics[width = 2cm]{figures/expert-18.png}};
  \node[draw,blue,thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 2cm,label=below:Canvas: render$(T)$] (canvas) at (0,-4) {};
  \draw[lightgray,ultra thin,step = 0.125] ([xshift = 0.5,yshift = 0.5]canvas.south west) grid ([xshift = -0.5,yshift = -0.5]canvas.north east);
  % draw partial image on canvas
  \draw (0.375cm,-3.25cm) circle (0.125cm);
  \draw (0.625cm,-3.625) circle (0.125cm);
  \draw (0.75cm,-3.375) circle (0.125cm);

@@

  \node(next)[draw,very thick,fit = (l1) (l2) (l3), dashed, label = below:{Next drawing command}] {};

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the trace set of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\citep{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[t]
  \caption{Primitive drawing commands currently supported by our model.}
\label{drawingCommandTable}
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes drawing command inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\end{table}

%% For the model in Fig.~\ref{architecture}, the distribution over the next drawing command factorizes as:
%% \begin{equation}
%%   \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta \left[t_k | a_\theta \left(f_\theta(I,\text{render}(T)) | \{t_j\}_{j = 1}^{k - 1}\right), \{t_j\}_{j = 1}^{k - 1}\right]
%% \end{equation}
%% where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
%% the target image, $T$ is a trace set, $\theta$ are the
%% parameters of the neural network, $f_\theta(\cdot,\cdot)$ is the
%% image feature extractor (convolutional network), and $a_\theta(\cdot|\cdot)$ is an attention mechanism. The distribution over
%% traces factorizes as:
%% \begin{equation}
%%   \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
%% \end{equation}
%% where $|T|$ is the length of trace $T$, the subscripts
%% on $T$ index drawing commands within the trace (so $T_n$ is a sequence of tokens: $t_1t_2\cdots t_K$), and the \verb|STOP|
%% token is emitted by the network to signal that the trace
%% explains the image.

We train the network by sampling trace sets $T$ and target
images $I$ for randomly generated scenes
and maximizing
%Eq.~\ref{objective}
the likelihood of $T$ given $I$ with respect to
the model parameters, $\theta$,
  by gradient ascent.
We trained the network on $10^5$ scenes, which takes a little less than a day on an Nvidia TitanX GPU.

Training does not require backpropagation across the sequence
of drawing commands: drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.  
This approach is like an autoregressive version of
Attend-Infer-Repeat~\citep{eslami1603attend}, but critically, without a learned
  recurrent memory state, e.g. an RNN.  We can do without an RNN
because we learn from ground truth (image, trace) pairs.  This
allows us to handle scenes with many objects without
worrying about the conditioning of gradients as they propagate
over long sequences, and makes training more straightforward: it
is just maximum likelihood estimation of the model parameters.

Our network can ``derender'' random synthetic images
%those shown in Fig.~\ref{trainingData}.
by doing a beam search decoding to
recover  trace sets traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
drawing command), it has no way of recovering from the error.  
To derender an image with $n$ objects, it must correctly predict $n$
commands -- so its probability of success will decrease
exponentially in $n$, assuming it has a nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\citep{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate
\clearpage\begin{wrapfigure}{R}{8cm}
  \centering
      \includegraphics[width = 7cm]{figures/editDistance36.png}
%    \includegraphics[width = 4cm]{figures/accuracy.png}
      \caption{Parsing~\LaTeX~output after training on diagrams with $\leq 12$ objects. It generalizes to scenes with many more objects. Neither the stochastic search nor the neural network are sufficient on their own. \# particles varies by model: we compare the models \emph{with equal runtime} ($\approx 1$ sec/object)}\label{syntheticResults}
\end{wrapfigure}
\noindent
for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathbb{R}$ uses the distance between two images as a proxy for a
likelihood. Unconventionally, the target distribution of the SMC sampler
includes the likelihood under the proposal distribution.
Intuitively, both the proposal
distribution and the distance function offer complementary signals for
whether a drawing command is correct.

Fig.~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We compare with a baseline that models the problem as image captioning (LSTM in Fig.~\ref{syntheticResults}).
Given the target image, this baseline produces the trace in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

%% \begin{figure}[H]\centering
%%   \begin{minipage}{6cm}
%%     \begin{minipage}[t]{3cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-58-7.png}}\end{minipage}%
%%     \begin{minipage}[t]{3cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-33-5.png}}\end{minipage}
%%     \caption{Network is trained to infer traces for randomly generated scenes like the two shown above. See supplement for details of the training data generation.}\label{trainingData}
%%   \end{minipage}\hfill%
%%   \begin{minipage}{7.5cm}\centering
%%     \includegraphics[width = 7cm]{figures/editDistance36.png}
%% %    \includegraphics[width = 4cm]{figures/accuracy.png}
%%       \caption{Parsing~\LaTeX~output after training on diagrams with $\leq 12$ objects. It generalizes to scenes with many more objects. Neither the stochastic search nor the neural network are sufficient on their own. \# particles varies by model: we compare the models \emph{with equal runtime} ($\approx 1$ sec/object)}\label{syntheticResults}
%%     \end{minipage}

%%   \end{figure}

\subsection{Generalizing to hand drawings}\label{generalizingTheHandDrawings}
\begin{wrapfigure}{R}{6cm}
\centering  \begin{minipage}[t]{2.5cm}\centering\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}\\
   \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): hand drawing. (b): Rendering of the trace our model infers for (a). (c): noisy rendering of (b).}\label{handDrawingExamples}
\end{wrapfigure}

A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Fig.~\ref{handDrawingExamples}; see supplement for details).

Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
  \begin{equation}
    -\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
    \end{equation}

Pixel-wise
distance
metrics are sensitive to the details of how
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details.
So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We evaluate our system on 100 hand-drawn figures; see Fig.~\ref{drawingSuccesses}--\ref{drawingFailures}.
These were drawn carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid, 
we made the drawings on graph paper.
For each drawing we annotated a ground truth trace and have the neurally guided SMC sampler
produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
We will show that the program synthesizer
corrects some of these small errors (Sec.~\ref{synthesizerHelpsParsing}).

%% Because the model sometimes makes mistakes % on hand drawings,
%% we envision it working as follows:
%% a user sketches a diagram,
%% and the system responds by proposing a few candidate interpretations.
%% The user could then select the one closest to their intention and edit it if necessary.

\setlength\tabcolsep{3.5pt}
\begin{figure}[H]
  \begin{minipage}[t]{8.25cm}
    \begin{tabular}{llll}
      \includegraphics[width = 2cm]{figures/expert-38.png}&
      \includegraphics[width = 2cm]{figures/expert-73.png}&
      \includegraphics[width = 2cm]{figures/expert-77.png}&
      \includegraphics[width = 2cm]{figures/expert-21.png}
      \\
      \includegraphics[width = 2cm]{figures/38-parse.png}     &
      \includegraphics[width = 2cm]{figures/73-parse.png}     &
      \includegraphics[width = 2cm]{figures/77-parse.png}    &
      \includegraphics[width = 2cm]{figures/21-parse.png}    
      \end{tabular}
    \caption{Left to right: Ising model, recurrent network architecture, figure from a deep learning textbook \cite{Goodfellow-et-al-2016}, graphical model}\label{drawingSuccesses}
  \end{minipage}
\hfill  \begin{minipage}[t]{4.25cm}
    \begin{tabular}{ll}
          \includegraphics[width = 2cm]{figures/expert-1.png}&
          \includegraphics[width = 2cm]{figures/expert-34.png}\\
      \includegraphics[width = 2cm]{figures/expert-1-parse.png}     &
              \includegraphics[width = 2cm]{figures/34-parse.png}    
    \end{tabular}

    \caption{Near misses. Rightmost: illusory contours (note: no SMC)}\label{drawingFailures}
  \end{minipage}
